The Splashback Radius of Planck-SZ clusters

### 2018-11-15 Paper_new/RSP_paper_new.tex
Dominiks changes

Given that the splashback radius represents a true halo boundary,
the observations of the splashback radius can be used to study a
variety of galaxy formation questions. Questions regarding the
timescales and the spatial scales within which star forming galaxies
quench after they fall into the cluster potential are of particular
interest to understand the fate of star formation in satellite
galaxies. In particular, if star forming galaxies quench before they
reach the apocenters of their orbit after infall, then they are not
expected to show a splashback feature in their density distribution.
In pursuit of this question, we also explore the dependence of the
cluster-galaxy cross-correlations separately for star forming and
quenched galaxy populations separated by color.

This paper is organized as follows. We introduce the observational
data sets we use in Section~\ref{sec:data}, namely the cluster and the
galaxy catalogs. We describe the methods and analysis procedures we
use in Section~\ref{sec:estimators}. We present and discuss our
results in Section~\ref{sec:Results}. Finally, we summarize our
findings in Section~\ref{sec:Conclusions} and discuss possible future
directions. Throughout the paper, we use a flat $\Lambda$CDM cosmology
with $\Omega_{\rm m}=0.27$ to convert redshifts and angles into
cosmological distances. Also, we denote three-dimensional distances by
$r$ and projected distances by $R$.

@@

\subsection{Cluster catalog}
\label{sec:clusters}

@@

We restrict ourselves to the redshift range $0.03 \leq z \leq 0.33$ in
order to have a similar redshift range used in
\citet{more2016detection}. Due to the larger beam size of the Planck
satellite, the cluster positions as reported in the catalog may be
mis-centered from the true centers of the galaxy clusters. We perform
a visual inspection of Pan-STARRS images taken around the detected
galaxy clusters in order to locate the nearest, brightest cluster
galaxy (BCG). We regard the position of the BCG as the true cluster
location, under the assumption that the BCG is located at the true,
gravitational center. Due to this crude assumption, we additionally 
study the effects of a possible, remaining mis-centering of the 
cluster positions in our model for the two-dimensional correlation 
function (see Appendix~\ref{sec:mis-centering}).

The final sample that we use consists of 596
galaxy clusters and is about an order of magnitude smaller compared
to the sample used in \citet{more2016detection}. The sample we use
in this paper has an average redshift of 0.177, and an average
cluster mass $M_{500c}$ of about $3.0 \cdot 10^{14}
h^{-1}$M$_{\sun}$. In Table~\ref{tab:cluster_catalogs} we compare
the main properties of the cluster catalog used in this work to that
used by \citet{more2016detection}. A map of the sky positions of the clusters
can be found in Figure~\ref{fig:planck_fig} in
Appendix~\ref{sec:figures}.

@@

In order to compute the color cut to be used, we first match spectroscopic
galaxies in the Sloan digital Sky Survey (DR 13) to their
Pan-STARRS photometry. We bin these galaxies in narrow redshift bins and
produce a histogram of the $g_{\mathrm{P1}}-r_{\mathrm{P1}}$ colors of the
galaxies in each bin. Due to the presence of the $4000\AA$-break present in the
quenched galaxy population, the two galaxy populations separate in such histograms
into two populations and we fit a double Gaussian distribution to it.
Based on the fitted distribution we use a cut in color to exclude a
contamination to the star-forming galaxy population with a confidence
of 3$\sigma$. We repeat this procedure for each redshift bin, to obtain a redshift
dependent color cut that separates the red from the blue population minimizing
the contamination of the blue population by red galaxies. Although being much
simpler and faster than calculating the individual k-corrections for each
galaxy, this procedure has its own shortcomings as discussed in
Appendix~\ref{sec:Errors}, in particular due to the photometric errors of red
galaxies. We would therefore exercise appropriate caution while interpreting
the results obtained by dividing galaxies in to color bins.

In our analysis, the same galaxy may be considered red or blue depending upon
the cluster redshift under consideration. Our method avoids the use of
uncertain photometric redshifts to derive k-corrections
\citep[cf.][]{baxter2017halo}. We will study the cross-correlations to derive
the splashback radii for these galaxy populations separately, as well as the
fractional contributions of the red and the blue galaxies to the total density
profile as a function of radius. The caveats in the interpretation of these
results are mentioned in the Appendix~\ref{sec:Errors}.

@@

curves show the corresponding first and second derivatives of the one-halo
term (namely $\rho_{\mathrm{in}}(r)f_{\mathrm{trans}}(r)$ in Equation~\ref{eq:dk14})

@@

Following \citet{baxter2017halo}, we also present the values of the
logarithmic derivatives at the location of the steepest slope for
the total three-dimensional cross-correlations, as well as those for
the inner halo term in the left and right hand panels of
Figure~\ref{fig:derivatives}, respectively. The logarithmic slope of the 
cross-correlation is significantly steeper than $-3$ at the location 
of the splashback radius, making it difficult to
be reproduced by classical fitting functions like the NFW profile,
which reach such slopes only asymptotically and even that only
without the presence of the outer 2-halo term. This provides
evidence for the existence of the splashback feature. From the second
derivative profiles we also note that the drop in the cross-correlation
signal is very sharp and happens within a factor 2 in radius.

@@

Next we present our measurements of the projected and
three-dimensional cross-correlations of the red and blue galaxy populations
satisfying $M_{i}-5\log h<-18.94$ with our SZ-selected cluster
sample. The projected cross-correlations are presented in Figure~\ref{fig:color_curve_2D},
whereas the three-dimensional counterparts are shown in 
Figure~\ref{fig:color_curve_3D}. The shaded regions show
the 68\% confidence intervals from our fits. The vertical, shaded
bands with different colors indicate the 68\% confidence regions of
the locations of the steepest slope of the two- and the
three-dimensional cross-correlations. The black dashed lines show
these ranges for the entire galaxy sample without regard to color.
The best fit parameters as well as the inferred splashback radii are
listed in Table~\ref{tab:fit_parameters}. We
see that the red galaxies have a steeper cross-correlation profile
than the blue galaxies. Although there seems to be  a tendency for the red
galaxies to have a smaller splashback radius, the differences we see
are not statistically significant given the current errors. The
slopes of the three-dimensional cross-correlations reach values
steeper than -2.5 at the splashback radius, which is difficult to model with
traditional NFW profiles. We take the detection of those steepening features 
as evidence for the existence of the splashback feature in both galaxy populations.
Taken at face value, our finding of the splashback feature in the blue
population would imply that there needs to be a reasonable fraction of
blue galaxies that fall into the cluster and continue to stay blue
even after reaching their apocenters.

@@

We caution however that the cut we use to define our samples was
defined based on spectroscopic galaxies, which tend to be brighter
than the typical galaxies we use for the cross-correlation. As we use
galaxies at fainter magnitudes for the cross-correlation, the
photometric errors increase dramatically and can scatter many more of
the red galaxy population into the blue galaxy sample. We report on
our investigations in Appendix~\ref{sec:Errors} as a cautionary tale,
and to guide future efforts to establish the splashback radius in the
blue galaxy population. We show that given the current state of the
data we cannot exclude a contamination of the correlation function of
blue galaxies from the red galaxy population and appropriate caution
be warranted in the interpretation or use of the results derived using
the color separated galaxy populations.

@@

The separation of the blue and red galaxy populations as described in
Section~\ref{sec:Color} is prone to photometric errors as explored in
the Appendix. A deeper galaxy catalog would be required in order to
confirm or rule out the existence of the splashback feature in the
blue galaxy population. The ongoing deep galaxy surveys such as the
Hyper Suprime-Cam (HSC) and the Dark Energy Survey (DES) would be able
to provide such data sets, albeit in a limited area. The Large
Synoptic Survey telescope (LSST) would eventually provide deep as well
as wide galaxy catalogs to eventually establish the locations of the
splashback radius at high significance.

@@

\begin{figure*}
    \includegraphics[width= \textwidth]{color_vs_z.pdf}
\caption{Scatter plot showing the $(g_{\mathrm{PS}}-r_{\mathrm{PS}})$ colors
of the matched galaxies versus their redshifts, where the red dots correspond
to galaxies identified as red from the SDSS spectroscopic colors and blue dots
correspond to galaxies identified as blue, respectively. The black, solid line
indicates the spline cut separating the two populations and excluding the red 
galaxies from the blue population at a confidence level of $3 \sigma$.}
   \label{fig:g-r_vs_z} 
\end{figure*}

To infer the color cut which separates star forming (blue) galaxies from the
quenched (red) galaxy population, we cross-matched the SDSS spectroscopic
sample to its Pan-STARRS photometry. In Figure~\ref{fig:g-r_vs_z}, we show the
plot of the Pan-STARRS $(g_{\mathrm{PS}}-r_{\mathrm{PS}})$ color as a 
function of redshift, and the color cut we use in our analysis. 
The SDSS spectroscopic sample is however quite shallow, 
and thus consists of galaxies that are brighter than the Pan-STARRS 
galaxies that we wish to cross-correlate. At fainter magnitudes, we can 
expect red galaxies to scatter into the blue population 
due to photometric errors, potentially contaminating the
correlation function measurement. This could erroneously cause the
splashback feature observed in the blue population. Therefore, we
need to assess the possiblity of such a contamination.

@@

In Figure~\ref{fig:mag_errors}, we show the photometric errors of the Pan-STARRS
galaxies as a function of magnitude in each of the bands we use. The
galaxies that we use in our cross-correlation analysis for the red/blue
galaxy population have an upper absolute magnitude limit of $M_{i}-5\log h=-18.94$. This
corresponds to a different apparent magnitude limit at each redshift.
For each spectroscopically matched galaxy, we figure out how faint the
apparent magnitude limit in the $i_{\mathrm{PS}}$ band is compared to the actual
apparent magnitude of the galaxy, recorded by Pan-STARRS. Assuming that the populations of
galaxies that we use to cross-correlate with clusters have the same
intrinsic colors, but are just fainter, we infer the true
intrinsic magnitude of the galaxies that we cross-correlate in the
$g_{\mathrm{PS}}$ and the $r_{\mathrm{PS}}$ band. We randomly perturb these magnitudes 
by the expected photometric errors at that magnitude as reported in Figure~\ref{fig:mag_errors}.

Next, we compute the fraction of galaxies which were intrinsically red
that now entered the blue population by the perturbation, where we
defined the separation of the populations using the same color
cut. We find such a contamination of intrinsically red galaxies to
the blue population to be about 5 \%. This value would apply for a galaxy population
measured in the field. Given that the red fraction is higher in clusters ($\sim
60$ \% in the whole cluster) than in the field ($\sim 40$ \%),
we expect there to be a larger proportion of red galaxies in the cluster which could
potentially contaminate the blue galaxies due to photometric errors.
This would roughly double the contamination to be about 10 \%.

This is a conservative estimate of the contamination, because the
spectroscopic galaxy sample we use has a large incompleteness at the
blue end at redshifts beyond $0.2$, where SDSS targetted the luminous
red galaxy population, recording only very few blue galaxies. 
Thus we are missing a lot of the blue galaxy
population in the highest redshift bins in our spectroscopic sample.
Nevertheless we assess if a $\sim 10$ \% contamination could
cause the splashback signal found in the cross-correlation for the blue galaxies.

>>>>

Given that the splashback radius represents a true halo boundary,
the observations of the splashback radius can be used to study a
variety of galaxy formation questions. Questions regarding the
timescales and the spatial scales within which star forming galaxies
quench after they fall into the cluster potential are of particular
interest to understand the fate of star formation in satellite
galaxies. In particular, if star forming galaxies quench before they
reach the apocenters of their orbit after infall, then they are not
expected to show a splashback feature in their density distribution.
In pursuit of this question, we also explore the dependence of the
cluster-galaxy cross-correlations separately for star forming and
quenched galaxy populations as separated by their color.

This paper is organized as follows. We introduce the observational
data sets we use in Section~\ref{sec:data}, namely the cluster and the
galaxy catalogs. We describe the methods and analysis procedures we
use in Section~\ref{sec:estimators}. We present and discuss our
results in Section~\ref{sec:Results}. Finally, we summarize our
findings in Section~\ref{sec:Conclusions} and discuss possible future
directions. Throughout the paper, we use a flat $\Lambda$CDM cosmology
with $\Omega_{\rm m}=0.27$ and a dimensionless Hubble parameter of $h=0.7$ 
to convert redshifts and angles into
cosmological distances. Also, we denote three-dimensional distances by
$r$ and projected distances by $R$.

@@

\subsection{Cluster catalog}
\label{sec:clusters}

\begin{figure}
    \includegraphics[scale=0.55]{Planck_summary.pdf}
\caption{Visualization of the distribution of the masses and redshifts of the used
clusters from the PSZ2 catalog.}
   \label{fig:placnk_summary}
\end{figure}

@@

We restrict ourselves to the redshift range $0.03 \leq z \leq 0.33$ in
order to have a similar redshift range as used in
\citet{more2016detection}. Due to the larger beam size of the Planck
satellite, the cluster positions as reported in the catalog may be
mis-centered from the true centers of the galaxy clusters. We perform
a visual inspection of Pan-STARRS images taken around the detected
galaxy clusters in order to locate the nearest, brightest cluster
galaxy (BCG). We regard the position of the BCG as the true cluster
location, under the assumption that the BCG is located at the true,
gravitational center. Due to this crude assumption, we additionally 
study the effects of a possible, remaining mis-centering of the 
cluster positions in our model for the two-dimensional correlation 
function (see Appendix~\ref{sec:mis-centering}).

The final sample that we use consists of 596
galaxy clusters and is about an order of magnitude smaller compared
to the sample used in \citet{more2016detection}. The sample we use
in this paper has an average redshift of 0.177, and an average
cluster mass $M_{500c}$ of about $3.0 \cdot 10^{14}
h^{-1}$M$_{\sun}$. In Table~\ref{tab:cluster_catalogs} we compare
the main properties of the cluster catalog used in this work to that
used by \citet{more2016detection}. Additionaly, Figure~\ref{fig:placnk_summary}
visualizes the distribution of the masses and redshifts of the used clusters.
A map of the sky positions of the clusters
can be found in Figure~\ref{fig:planck_fig} in
Appendix~\ref{sec:figures}.

@@

In order to compute the color cut to be used, we first match spectroscopic
galaxies in the Sloan Digital Sky Survey (DR 8) to their
Pan-STARRS photometry. We bin these galaxies in narrow redshift bins and
produce a histogram of the $g_{\mathrm{P1}}-r_{\mathrm{P1}}$ colors of the
galaxies in each bin. Due to the presence of the $4000\AA$-break present in the
quenched galaxy population, the two galaxy populations separate in such histograms
into two populations and we fit a double Gaussian distribution to it.
Based on the fitted distribution we use a cut in color to exclude a
contamination to the star-forming galaxy population with a confidence
of 3$\sigma$. We repeat this procedure for each redshift bin, to obtain a redshift
dependent color cut that separates the red from the blue population minimizing
the contamination of the blue population by red galaxies. Although being much
simpler and faster than calculating the individual k-corrections for each
galaxy, this procedure has its own shortcomings as discussed in
Appendix~\ref{sec:Errors}, in particular due to the photometric errors of red
galaxies. We would therefore exercise appropriate caution while interpreting
the results obtained by dividing galaxies into color bins.

In our analysis, the same galaxy may be considered red or blue depending upon
the cluster redshift under consideration. Our method avoids the use of
uncertain photometric redshifts to derive k-corrections
\citep[cf.][]{baxter2017halo}. We will study the cross-correlations to derive
the splashback radii for these galaxy populations separately. 
The caveats in the interpretation of these
results are mentioned in the Appendix~\ref{sec:Errors}.

@@

curves show the corresponding first and second order derivatives of the one-halo
term (namely $\rho_{\mathrm{in}}(r)f_{\mathrm{trans}}(r)$ in Equation~\ref{eq:dk14}).

@@

Following \citet{baxter2017halo}, we also present the values of the
first and second order logarithmic derivatives at the location of the steepest slope for
the total three-dimensional cross-correlations, as well as those for
the inner halo term in the middle and right hand panels of
Figure~\ref{fig:derivatives}, respectively. The logarithmic slope of the 
cross-correlation is significantly steeper than $-3$ at the location 
of the splashback radius, making it difficult to
be reproduced by classical fitting functions like the NFW profile,
which reach such slopes only asymptotically and even that only
without the presence of the outer 2-halo term. This provides
evidence for the existence of the splashback feature. From the second
order derivative profiles we also note that the drop in the cross-correlation
signal is very sharp and happens within a factor 2 in radius.

@@

Next, we present our measurements of the projected and
three-dimensional cross-correlations of the red and blue galaxy populations
satisfying $M_{i}-5\log h<-18.94$ with our SZ-selected cluster
sample. The projected cross-correlations are presented in Figure~\ref{fig:color_curve_2D},
whereas the three-dimensional counterparts are shown in 
Figure~\ref{fig:color_curve_3D}. The shaded regions show
the 68\% confidence intervals from our fits. The vertical, shaded
bands with different colors indicate the 68\% confidence regions of
the locations of the steepest slope of the two- and the
three-dimensional cross-correlations. The black dashed lines show
these ranges for the entire galaxy sample without regard to color.
The best fit parameters as well as the inferred splashback radii are
listed in Table~\ref{tab:fit_parameters}. We
see that the red galaxies have a steeper cross-correlation profile
than the blue galaxies. Although there seems to be  a tendency for the red
galaxies to have a smaller splashback radius, the differences we see
are not statistically significant given the current errors. The
slopes of the three-dimensional cross-correlations reach values
steeper than -2.5 at the splashback radius, which is difficult to model with
traditional NFW profiles. We take the detection of those steepening features 
as evidence for the existence of the splashback feature in both galaxy populations.
Taken at face value, our finding of the splashback feature in the blue
population would imply that there needs to be a reasonable fraction of
blue galaxies that fall into the cluster and continue to stay blue
even after reaching their apocenters.

@@

We caution however that the cut we use to define our samples was
defined based on spectroscopic galaxies, which tend to be brighter
than the typical galaxies we use for the cross-correlation. As we use
galaxies at fainter magnitudes for the cross-correlation, the
photometric errors increase dramatically and can scatter many more of
the red galaxies into the blue galaxy sample. We report on
our investigations in Appendix~\ref{sec:Errors} as a cautionary tale,
and to guide future efforts to establish the splashback radius in the
blue galaxy population. We show that given the current state of the
data we cannot exclude a contamination of the correlation function of
blue galaxies from the red galaxy population and appropriate caution
be warranted in the interpretation or use of the results derived using
the color-separated galaxy populations.

@@

The separation of the blue and red galaxy populations as described in
Section~\ref{sec:Color} is prone to photometric errors as explored in
 Appendix~\ref{sec:Errors}. A deeper galaxy catalog would be required in order to
confirm or rule out the existence of the splashback feature in the
blue galaxy population. The ongoing deep galaxy surveys such as the
Hyper Suprime-Cam (HSC) and the Dark Energy Survey (DES) would be able
to provide such data sets, albeit in a limited area. The Large
Synoptic Survey telescope (LSST) would eventually provide deep as well
as wide galaxy catalogs to eventually establish the locations of the
splashback radius at high significance.

@@

\begin{figure*}
    \includegraphics[width= \textwidth]{color_vs_z.pdf}
\caption{Scatter plot showing the $(g_{\mathrm{P1}}-r_{\mathrm{P1}})$ colors
of the matched galaxies versus their redshifts, where the red dots correspond
to galaxies identified as red from the SDSS spectroscopic colors and blue dots
correspond to galaxies identified as blue, respectively. The black, solid line
indicates the spline cut separating the two populations and excluding the red 
galaxies from the blue population at a confidence level of $3 \sigma$.}
   \label{fig:g-r_vs_z} 
\end{figure*}

To infer the color cut which separates star forming (blue) galaxies from the
quenched (red) galaxy population, we cross-matched the SDSS spectroscopic
sample to its Pan-STARRS photometry. In Figure~\ref{fig:g-r_vs_z}, we show the
plot of the Pan-STARRS $(g_{\mathrm{P1}}-r_{\mathrm{P1}})$ color as a 
function of redshift, and the color cut we use in our analysis. 
The SDSS spectroscopic sample is however quite shallow, 
and thus consists of galaxies that are brighter than the Pan-STARRS 
galaxies that we wish to cross-correlate. At fainter magnitudes, we can 
expect red galaxies to scatter into the blue population 
due to photometric errors, potentially contaminating the
correlation function measurement. This could erroneously cause the
splashback feature observed in the blue population. Therefore, we
need to assess the possiblity of such a contamination.

@@

In Figure~\ref{fig:mag_errors}, we show the photometric errors of the Pan-STARRS
galaxies as a function of magnitude in each of the bands we use. The
galaxies that we use in our cross-correlation analysis for the red/blue
galaxy population have an upper absolute magnitude limit of $M_{i}-5\log h=-18.94$. This
corresponds to a different apparent magnitude limit at each redshift.
For each spectroscopically matched galaxy, we figure out how faint the
apparent magnitude limit in the $i_{\mathrm{P1}}$ band is compared to the actual
apparent magnitude of the galaxy, recorded by Pan-STARRS. Assuming that the populations of
galaxies that we use to cross-correlate with clusters have the same
intrinsic colors, but are just fainter, we infer the true
intrinsic magnitude of the galaxies that we cross-correlate in the
$g_{\mathrm{P1}}$ and the $r_{\mathrm{P1}}$ band. We randomly perturb these magnitudes 
by the expected photometric errors at that magnitude as reported in Figure~\ref{fig:mag_errors}.

Next, we compute the fraction of galaxies which were intrinsically red
that now entered the blue population by the perturbation, where we
defined the separation of the populations using the same color
cut. We find such a contamination of intrinsically red galaxies to
the blue population to be about 5\%. This value would apply for a galaxy population
measured in the field. Given that the red fraction is higher in clusters ($\sim
60$\% in the whole cluster) than in the field ($\sim 40$\%),
we expect there to be a larger proportion of red galaxies in the cluster which could
potentially contaminate the blue galaxies due to photometric errors.
This would roughly double the contamination to be about 10\%.

This is a conservative estimate of the contamination, because the
spectroscopic galaxy sample we use has a large incompleteness at the
blue end at redshifts beyond $0.2$, where SDSS targetted the luminous
red galaxy population, recording only very few blue galaxies. 
Thus we are missing a lot of the blue galaxy
population in the highest redshift bins in our spectroscopic sample.
Nevertheless we assess if a $\sim 10$\% contamination could
cause the splashback signal found in the cross-correlation for the blue galaxies.
